On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-04 model.tex
minor edits at home

\begin{align}\label{vshear_split}
  \mathcal{V} =  \frac{\left|r\p_z\Omega\right|_{\text{
        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
        temp. gradient}}} \sim
  \epsilon \frac{\mathrm{max}\left(\delta^2,
    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
\end{align}
where $\delta\equiv \Htilde/\Hgas$. 
Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
vertical shear due to variations in the dust-to-gas ratio can be at
most comparable to that due to the radial temperature gradient; and
this only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so
that $\mathcal{V}\sim 1/|q|$). Otherwise, vertical shear is
predominantly due to $\p_rc_s^2$.

\subsection{Vertical buoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thin dust layers such that $H_\epsilon \ll Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

We see that $N_z^2\sim \epsilon \OmK^2$, while $r\p_z\Omega^2\sim
h_\mathrm{g}\OmK^2$ for a thin disk where vertical shear is mainly due to the
radial temperature gradient. Since buoyancy is stabilizing, we expect
the dust-induced buoyancy forces to stabilize the disk against the VSI
where $\epsilon \gtrsim h_\mathrm{g}$.

>>>>

\begin{align}\label{vshear_split}
\frac{\left|r\p_z\Omega\right|_{\text{
        dust/gas gradient}}}{\left|r\p_z\Omega\right|_{\text{
        temp. gradient}}} \sim
  \epsilon \frac{\mathrm{max}\left(\delta^2,
    1\right)}{\left|q\right|\left(\epsilon + \delta^2\right)},
\end{align}
where $\delta\equiv \Htilde/\Hgas$. 
Since $|q|=O(1)$ in PPDs, Eq. \ref{vshear_split} indicates that
vertical shear due to variations in the dust-to-gas ratio can be at
most comparable to that due to the radial temperature gradient. 
This only occurs for thin dust layers with $\delta^2\ll \epsilon$ (so 
the above ratio $\sim 1/|q|$). Otherwise, vertical shear is associated
with $\p_rc_s^2$.

\subsection{Vertical buoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

### 2016-08-04 result.tex
minor edits at home

>>>>

We see that $N_z^2\sim \epsilon \OmK^2$, while $r\p_z\Omega^2\sim
h_\mathrm{g}\OmK^2$ for a thin disk where vertical shear is mainly due to the
radial temperature gradient. Since buoyancy is stabilizing, we expect
the dust-induced buoyancy forces to stabilize the disk against the VSI
where $\epsilon \gtrsim h_\mathrm{g}$.
